### 2024-02-14 outputs/paper.qmd
Added graphs to qmd.

---
title: "Information and Airbnb, An Analysis on the Impact of Reviews and Ratings on Airbnb Prices in Canada"
author: 
  - Alaina Hu and Ahad Qureshi
thanks: "Code and data are available at: https://github.com/alainahu/airbnb_analysis"
date: today
date-format: long
abstract: ''
format: pdf
toc: true
number-sections: true
#bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(knitr)
library(kableExtra)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(RColorBrewer)
library(gridExtra)
```
# Introduction

# Data

```{r}
#| echo: false
airbnb <- read_csv("~/airbnbtp2/inputs/data/airbnb_cleaned_data.csv")
```

```{r}
#| fig-cap: Distribution of Airbnb prices per night (including cleaning fees)
#| label: price_dist
#| echo: false

df1 <- airbnb
ggplot(df1, aes(x=new_price)) +
  geom_histogram(aes(y=..density..), binwidth=10, colour="black", fill="grey") + # binwidth adjusted for demonstration
  scale_x_continuous(limits=c(0, 500), breaks=seq(0, 500, by=50), name="Price") +
  scale_y_continuous(name="Density") +
  theme_minimal() +
  theme(plot.background = element_rect(fill="white"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust=0.5))

summary(df1$new_price)
```

```{r}
summary_df <- airbnb |>
  summarise(
    Mean_New_Price = mean(new_price, na.rm = TRUE),
    Median_New_Price = median(new_price, na.rm = TRUE),
    SD_New_Price = sd(new_price, na.rm = TRUE))
```

```{r}
```

```{r}
#| tbl-cap: Summary Statistics for the Reviews and Ratings 
#| label: summ_stats
#| echo: false

summary <- airbnb |>
  summarise(
    Mean_Review = mean(review, na.rm = TRUE),
    Median_Review = median(review, na.rm = TRUE),
    SD_Review = sd(review, na.rm = TRUE),
    Min_Review = min(review, na.rm = TRUE),
    Max_Review = max(review, na.rm = TRUE),
    Mean_Guest = mean(guest_satisfaction_overall, na.rm = TRUE),
    Median_Guest = median(guest_satisfaction_overall, na.rm = TRUE),
    SD_Guest = sd(guest_satisfaction_overall, na.rm = TRUE),
    Min_Guest = min(guest_satisfaction_overall, na.rm = TRUE),
    Max_Guest = max(guest_satisfaction_overall, na.rm = TRUE),
    Mean_Accuracy = mean(accuracy_rating, na.rm = TRUE),
    Median_Accuracy = median(accuracy_rating, na.rm = TRUE),
    SD_Accuracy = sd(accuracy_rating, na.rm = TRUE),
    Min_Accuracy = min(accuracy_rating, na.rm = TRUE),
    Max_Accuracy = max(accuracy_rating, na.rm = TRUE),
    Mean_Clean = mean(cleanliness_rating, na.rm = TRUE),
    Median_Clean = median(cleanliness_rating, na.rm = TRUE),
    SD_Clean = sd(cleanliness_rating, na.rm = TRUE),
    Min_Clean = min(cleanliness_rating, na.rm = TRUE),
    Max_Clean = max(cleanliness_rating, na.rm = TRUE),
    Mean_Location = mean(location_rating, na.rm = TRUE),
    Median_Location = median(location_rating, na.rm = TRUE),
    SD_Location = sd(location_rating, na.rm = TRUE),
    Min_Location = min(location_rating, na.rm = TRUE),
    Max_Location = max(location_rating, na.rm = TRUE),
    Mean_Value = mean(value_rating, na.rm = TRUE),
    Median_Value = median(value_rating, na.rm = TRUE),
    SD_Value = sd(value_rating, na.rm = TRUE),
    Min_Value = min(value_rating, na.rm = TRUE),
    Max_Value = max(value_rating, na.rm = TRUE)
  )
long_summary_stats <- summary |>
  pivot_longer(cols = everything(), names_to = "Statistic", values_to = "Value") |>
  separate(Statistic, into = c("Measure", "Variable"), sep = "_") |>
  spread(key = Measure, value = Value) |>
  select(Variable, Mean, Median, SD, Min, Max)

long_summary_stats$Variable <- c("Accuracy Rating", "Cleanliness Rating", "Guest Overall Satisfaction", 
                                 "Location Rating", "Number of Reviews", "Value Rating")

# Now use kable to create the markdown table
kable(long_summary_stats, 
      format = "markdown", 
      digits = 2,
      col.names = c("Variable", "Mean", "Median", "Standard Deviation", "Min", "Max"),
      row.names = FALSE, # We already have the Variable column, so we don't want to print row names
      align = c('l', 'r', 'r', 'r', 'r', 'r'), # Align the columns left for variable name and right for numbers 
      )

```
```{r}
#| echo: false
set2Colors <- brewer.pal(8, "Set2")
toronto <- set2Colors[1] # Light green
vancouver <- set2Colors[2]
```

```{r}
#| tbl-cap: Plot of Listing Reviews and Price
#| label: review_price
#| echo: false
#| 
ggplot(airbnb, aes(x = review, y = new_price, color = city)) +
         geom_point(size = 0.75) +
         geom_smooth(aes(color = city), method = "lm", se = FALSE, size = 1.5) +        geom_smooth(method = "lm", color = "black", se = TRUE, size = 1) +  
         labs(
       x = "Number of Reviews", 
       y = "Price") +
  theme_minimal() +
  theme(plot.background = element_rect(color = "black", fill = "white"))
  scale_color_brewer(palette = "Set2")
```

```{r}
#| tbl-cap: Plot of Guest Overall Satisfaction Score and Price
#| label: guest_satisf
#| echo: false
#| 
ggplot(airbnb, aes(x = guest_satisfaction_overall, y = new_price, color = city)) +
         geom_point(size = 0.75) +
         geom_smooth(aes(color = city), method = "lm", se = FALSE, size = 1.5) +        geom_smooth(method = "lm", color = "black", se = TRUE, size = 1) +  
         labs(
       x = "Guest Overall Satisfaction Score", 
       y = "Price") +
  theme_minimal() +
  theme(plot.background = element_rect(color = "black", fill = "white")) 
  scale_color_manual(values = c("toronto" = toronto, "vancouver" = vancouver))
```

```{r}
unique_values <- unique(airbnb$accuracy_rating)

# Print the unique values
print(unique_values)
```

```{r}
#| tbl-cap: Plot of Accuracy Rating and Price
#| label: accuracy_price
#| echo: false

ggplot(airbnb, aes(x=factor(accuracy_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill="grey") +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Accuracy Rating", y = "Price") +
  theme_minimal()

```

```{r}
#| echo: false
toronto_data <- airbnb |>
  filter(city == "toronto")

toronto_accuracy <- ggplot(toronto_data, aes(x=factor(accuracy_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill=toronto) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Accuracy Rating", y = "Price") +
  theme_minimal()
```

```{r}
#| echo: false
vancouver_data <- airbnb |>
  filter(city == "vancouver")

vancouver_accuracy <- ggplot(vancouver_data, aes(x=factor(accuracy_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill=vancouver) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Accuracy Rating", y = "Price") +
  theme_minimal()
```

```{r}
#| tbl-cap: Plot of Accuracy Rating and Price in Toronto (left) and Vancouver (right)
#| label: accuracy_price_2
#| echo: false
grid.arrange(toronto_accuracy, vancouver_accuracy, ncol = 2)
```

```{r}
#| tbl-cap: Plot of Cleanliness Rating and Price
#| label: clean_price
#| echo: false

ggplot(airbnb, aes(x=factor(cleanliness_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill="grey") +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Cleanliness Rating", y = "Price") +
  theme_minimal()

```

```{r}
#| echo: false

toronto_cleanliness <- ggplot(toronto_data, aes(x=factor(cleanliness_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill=toronto) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Cleanliness Rating", y = "Price") +
  theme_minimal()
```

```{r}
#| echo: false

vancouver_cleanliness <- ggplot(vancouver_data, aes(x=factor(cleanliness_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill=vancouver) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Cleanliness Rating", y = "Price") +
  theme_minimal()
```

```{r}
#| tbl-cap: Plot of Cleanliness Rating and Price in Toronto (left) and Vancouver (right)
#| label: clean_price_2
#| echo: false
grid.arrange(toronto_cleanliness, vancouver_cleanliness, ncol = 2)
```

```{r}
#| tbl-cap: Plot of Location Rating and Price
#| label: location_price
#| echo: false

ggplot(airbnb, aes(x=factor(location_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill="grey") +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Location Rating", y = "Price") +
  theme_minimal()

```

```{r}
#| echo: false

toronto_location <- ggplot(toronto_data, aes(x=factor(location_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill=toronto) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Location Rating", y = "Price") +
  theme_minimal()
```

```{r}
#| echo: false

vancouver_location <- ggplot(vancouver_data, aes(x=factor(location_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill=vancouver) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Location Rating", y = "Price") +
  theme_minimal()
```

```{r}
#| tbl-cap: Plot of Location Rating and Price in Toronto (left) and Vancouver (right)
#| label: location_price_2
#| echo: false
grid.arrange(toronto_location, vancouver_location, ncol = 2)
```

```{r}
#| tbl-cap: Plot of Value Rating and Price
#| label: value_price
#| echo: false

ggplot(airbnb, aes(x=factor(value_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill="grey") +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Value Rating", y = "Price") +
  theme_minimal()

```

```{r}
#| echo: false

toronto_value <- ggplot(toronto_data, aes(x=factor(value_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill=toronto) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Value Rating", y = "Price") +
  theme_minimal()
```

```{r}
#| echo: false

vancouver_value <- ggplot(vancouver_data, aes(x=factor(value_rating), y=new_price)) +
  geom_violin(trim=FALSE, fill=vancouver) +
  stat_summary(fun=mean, geom="point", shape=23, size=2, fill="white") +
  scale_x_discrete(limits = as.character(0:10)) + 
  labs(x = "Value Rating", y = "Price") +
  theme_minimal()
```

```{r}
#| tbl-cap: Plot of Value Rating and Price in Toronto (left) and Vancouver (right)
#| label: value_price_2
#| echo: false
grid.arrange(toronto_value, vancouver_value, ncol = 2)
```